The Design and Implementation of the Synchronous Language \CEU

### 2016-03-31 paper.tex
...

To mitigate this threat, \CEU performs a temporal analysis at compile time and 
detects concurrent accesses to shared variables, as follows:
if a variable is written in a trail segment, then a concurrent trail segment 
cannot read or write to that variable, nor dereference a pointer of that 
variable type.
%An analogous policy is applied for pointers \emph{vs} variables and pointers 
%\emph{vs} pointers.
%
A trail segment is a sequence of statements followed by an \code{await} (or 
termination).
Concurrency in \CEU is characterized when two or more trail segments in 
parallel react to the same input event.
%
Considering the examples in Figure~\ref{lst.shared}:
%
\begin{itemize}
\item The assignments to \code{x} in lines 2 and 5 in \ax
      \textbf{cannot} be concurrent because they are not in parallel trails.
\item The assignments to \code{x} in lines 5 and 8 in \ax
      \textbf{cannot} be concurrent because they \textbf{cannot} execute during 
      the same reaction.
\item The assignments to \code{y} in lines 5 and 8 in \bx
      \textbf{can} be concurrent because they are in parallel trails and 
      \textbf{can} execute during the same reaction.
\end{itemize}
%
The algorithm for the analysis, which is depicted in 
Section~\ref{sec.impl.analysis}, inspects all possible \code{await} statements 
that precede a variable access and keep a list with all corresponding awaking 
events.
Then, it checks all accesses in parallel trails to see if they share an awaking 
event.
If it is the case, the compiler warns about the suspicious accesses.

@@

TODO: internal events, emits are propagated from preceding awaits to input

The compile-time \emph{temporal analysis} phase detects inconsistencies in \CEU 
programs.
Here, we focus on the algorithm that detects suspicious access to shared 
variables, as discussed in Section~\ref{sec.ceu.analysis}.

For each node representing a statement in the program AST, we keep the set of 
events $I$ (for \emph{incoming}) that can lead to the execution of the node, 
and also the set of events $O$ (for \emph{outgoing}) that can terminate the 
node.

A node inherits the set $I$ from its direct parent and calculates $O$ according 
to its type:
%
\begin{itemize}
%
\item Nodes that represent expressions, assignments, \C calls, and declarations 
simply reproduce $O=I$, as they do not await;
%
\item An \code{await e} statement has $O=\{e\}$.
%
\item A \code{break} statement has $O=\{\}$ as it escapes the innermost 
\code{loop} and never terminates, i.e., never proceeds to the statement 
immediately following it (see also \code{loop} below);
%
\item A \emph{sequence node (;)} modifies each of its children to have 
$I_n=O_{n-1}$.
The first child inherits $I$ from its parent node, and the set $O$ for the 
sequence node is copied from its last child, i.e., $O=O_n$.
%
\item A \code{loop} node includes its body's $O$ on its own $I$ ($I=I \cup 
O_{body}$), as the loop is also reached from its own body.
The union of all \code{break} statements' $O$ forms the set $O$ for a 
\code{loop}.
%
\item An \code{if} node has $O=O_{true} \cup O_{false}$.
%
\item A parallel composition may terminate from any of its branches, hence $O = 
O_1 \cup ... \cup O_n$.
\end{itemize}

>>>>

To mitigate this threat, \CEU performs a temporal analysis at compile time to 
detect concurrent accesses to shared variables:
if a variable is written in a trail segment, then a concurrent trail segment 
cannot read or write to that variable, nor dereference a pointer of that 
variable type.
%An analogous policy is applied for pointers \emph{vs} variables and pointers 
%\emph{vs} pointers.
%
A trail segment is a sequence of statements followed by an \code{await} (or 
termination).
Concurrency in \CEU is characterized when two or more trail segments in 
parallel react to the same input event.
%
Considering the examples in Figure~\ref{lst.shared}:
%
\begin{itemize}
\item The assignments to \code{x} in lines 2 and 5 in \ax
      \textbf{cannot} be concurrent because they are not in parallel trails.
\item The assignments to \code{x} in lines 5 and 8 in \ax
      \textbf{cannot} be concurrent because they \textbf{cannot} execute during 
      the same reaction.
\item The assignments to \code{y} in lines 5 and 8 in \bx
      \textbf{can} be concurrent because they are in parallel trails and 
      \textbf{can} execute during the same reaction.
\end{itemize}
%
The algorithm for the analysis, which is depicted in 
Section~\ref{sec.impl.analysis}, inspects all possible \code{await} statements 
that precede a variable access and keep a list with all corresponding awaking 
events.
Then, it checks all accesses in parallel trails to see if they share an awaking 
event.
If it is the case, the compiler warns about the suspicious accesses.

@@

The compile-time \emph{temporal analysis} phase detects inconsistencies in \CEU 
programs.
Here, we focus on the algorithm that detects suspicious concurrent statements, 
such as accesses to shared variables, as discussed in 
Section~\ref{sec.ceu.analysis}.

For each node representing a statement in the program AST, we keep the set of 
input events $I$ (for \emph{incoming}) that can lead to the execution of the 
node, and also the set of input events $O$ (for \emph{outgoing}) that can 
terminate the node.

A node inherits the set $I$ from its direct parent and calculates $O$ according 
to its type:
%
\begin{itemize}
%
\item Nodes that represent expressions, assignments, \C calls, and declarations 
simply reproduce $O=I$, as they do not await;
%
\item An \code{await E} statement has $O=\{E\}$, where \code{E} is an input 
event (see internal events below).
%
\item A \code{break} statement has $O=\{\}$ as it escapes the innermost 
\code{loop} and never terminates, i.e., never proceeds to the statement 
immediately following it (see also \code{loop} below);
%
\item A \emph{sequence node (;)} modifies each of its children to have 
$I_n=O_{n-1}$.
The first child inherits $I$ from its parent node, while the set $O$ for the 
whole node is copied from its last child, i.e., $O=O_n$.
%
\item A \code{loop} node includes its body's $O$ on its own $I$ ($I=I \cup 
O_{body}$), as the loop is also reached from its own body.
The union of all \code{break} statements' $O$ forms the set $O$ for a 
\code{loop}.
%
\item An \code{if} node has $O=O_{true} \cup O_{false}$.
%
\item A parallel composition may terminate from any of its branches, hence $O = 
O_1 \cup ... \cup O_n$.
\item For internal events, an \code{await} awakes from any input that leads to 
    any matching \code{emit} in a trail in parallel:
    \begin{itemize}
    \item An \code{emit e} terminates in the same reaction, having $O=I$.
    \item An \code{await e} has $O=I_{e1} \cup ... \cup I_{eN}$,
        where $e1...eN$ are \code{emit e} statements in trails in parallel.
    \end{itemize}
\end{itemize}
